# Getting started with the Twitter API in R

## Contents

1. Introduction
2. Keys and secrets
3. Retrieving tweets from your timeline
4. Search tweets about a topic


## 1. Introduction

### REST APIs 

The Twitter Platform is made up of a number of APIs and tools, some of which are the REST APIs.  If you're not sure what that means, Wikipedia tells us that:

“RESTful systems typically, but not always, communicate over Hypertext Transfer Protocol (HTTP) with the same HTTP verbs (GET, POST, PUT, DELETE, etc.) that web browsers use to retrieve web pages and to send data to remote servers.  REST systems interface with external systems as web resources identified by Uniform Resource Identifiers (URIs), for example /people/tom, which can be operated upon using standard verbs such as DELETE /people/tom.”

### Rate limiting

The Twitter APIs are rate-limited, meaning that in every 15 minute window, there is a maximum number of request that you can make with a single command.

For the GET commands, these are either 15 or 180, depending on the individual command.
The Search API can be used to query tweets; however, the API documentation notes that relevancy is prioritised over completeness, and so if one wishes to access every tweet with a certain match, a Streaming API may be more appropriate.

## 2. Keys and secrets

When you sign up to use the Twitter APIs, there are four different values which you need to use to be able to use the API within R: 

1. consumer key
2. consumer secret
3. token
4. token secret.

You can think of these as being akin to a username, password, a user ID for the particular app, and the password for that app to access the Twitter API.

If you wish to share your scripts with others, you should remove these values from your scripts to ensure account security.  A simple way to ensure privacy is to use the `gWidgets` package to prompt you with a dialog box to enter your keys and secrets when you run your script.

```{r eval=FALSE}
library(gWidgets)
options(guiToolkit="tcltk") 
consKey <- ginput("Enter your consumer key:")
consSecret<-ginput("Enter your consumer secret:")
token<-ginput("Enter your token:")
tokenSecret<-ginput("Enter your token secret:")
```

```{r echo=FALSE}
consKey="g06IsRZmEJd1lUp0iC5cnFBdl"
consSecret="fPBEZmA0OB2lCmKET6jG1NvI7QorhrE5Hcym7b9kz4NlbYRzqA"
token="317965625-Dtwle8hz9yYzicDmWNCWquG91tAPYLekxDsohCVo"
tokenSecret="09alcq8nB6Sl2EE9S1oYwTBoUxQkAXSpsOdYMspVwmAHM"

```

The `httr` package is useful for working with URLs and HTTP.  We start by taking our keys and secrets and creating objects which will allow us to interact with the API.

```{r}
library(httr)

# start the authorisation process
myapp = oauth_app("twitter", key=consKey, secret=consSecret)

# sign using token and token secret
sig = sign_oauth1.0(myapp, token=token, token_secret=tokenSecret)

```

## 3. Retrieving tweets from your timeline

One of the tasks you might wish to do it to retrieve tweets from your timeline.  

To accomplish this, use the `GET` command, supplying to it as parameters the below URL and the variable created earlier to sign.

```{r}
my_timeline=GET("https://api.twitter.com/1.1/statuses/home_timeline.json", sig)
```

The URL we input above related to part of the Twitter API.  Documentation for this command can be found [here](https://dev.twitter.com/rest/reference/get/statuses/home_timeline).

If you browse the sidebar of this site, you can also see other examples of things you can retrieve using GET.

### Examining the content

OK, so we've retrieved content from twitter, but what do we do with it now?

We can use the `content` function to get the JSON data as a structured R objects.  As this can be hard to read, we use the `jsonlite` package to reformat it as a data.frame.

```{r}
library(jsonlite)

json1 = content(my_timeline)
json2 = jsonlite::fromJSON(toJSON(json1))
```

Let's take a look at the data.  For reference purposes, the first 4 columns are as follows:

1. created_at - when the tweet was sent
2. id - ID as a numeric value
3. id_str - ID as a string
4. text - the content of the tweet

There are more columns than this, although for the purpose of simplicity, here we will just examine the first 4.

Here are the 3 most recent tweets on my timeline:

```{r}
json2[1:3,1:4]
```

### Retweets

Let's have a look at another example.  Here I use the API to look at which of my recent tweets have been retweeted by others.
```{r}
retweets=GET("https://api.twitter.com/1.1/statuses/retweets_of_me.json", sig)
json_rtw = content(retweets)
json_rtw2 = jsonlite::fromJSON(toJSON(json_rtw))
json_rtw2[1:5,4]
```

## 4. Search tweets about a topic

Finally, we may want to take a look at tweets about a certain topic.  Keep in mind my previous comments about the difference in results returned by the Search API and the Streaming API.

Let's have a look at the content of recent tweets about [Shiny](http://shiny.rstudio.com/)

```{r}
shiny_tweets=GET("https://api.twitter.com/1.1/search/tweets.json?q=rshiny", sig)
json_shiny = content(shiny_tweets)
json_shiny2 = jsonlite::fromJSON(toJSON(json_shiny))
statuses<-json_shiny2$statuses
statuses[1:10,4]
```
Thanks to the [Getting and Cleaning Data](https://www.coursera.org/learn/data-cleaning/home/welcome) course by John Hopkins University on Coursera for providing the basis for much of this content.

